---
title       : MT Cars Dataset Exploratory Analysis
subtitle    : Coursera Developing Data Products: Pitch Presentation Assignment
author      : Marco Lunardi (Please use arrow keys on your keyboard to move through slides, thank you)
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## The MTCars Dataset

The MTCars Dataset is available into the "datasets" package in R.
At the R console prompt you can type "library(datasets)", and the "? mtcars" to see a brief description about it.

The dataset gathers data about 32 cars from a 1974 issue of Motor Trend Magazine (quite old data actually), showing values for 11 variables.

The main aim of the dataset is to analyze the influence on the efficiency of the cars by some relevant variables (weight, trasmission, engine size, acceleration, etc...),
in particular which type of transmission (manual or automatic) was the best in terms of fuel saving.

The measure adopted to evaluate the efficiency is MPG (= Miles per US Gallon).
So, the more efficient cars have a higher MPG and then lower fuel consumption.


---

## My Shiny App

My App is designed to perform a very basic exploratory analysis on mtcars data, in order to get a first view
about the interactions among the most relevant variables within the mtcars dataset.

You can find my app at the following web address:  https://marcolunardi.shinyapps.io/DDP01APP/

On the upper left side of the app page you can set the values for two vary important variables:
1. The engine size, through the cylinders number: from 4 to 8
2. The transmission type: manual or automatic

Once you choose the two values, the app dinamically changes the content on the right side of the app page.
1. It changes the content of the title, reflecting your choices.
2. It computes the average weight and MPG for the cars that correspond to your criteria.
3. It plots a scatterplot graph in which each point represents just the cars you selected through your settings (red points if "manual transmission", black points if "automatic").

The following slide computes the averages of weights and MPGs values, and plots a scatterplot, for all the cars in the mtcars dataset.
That are the reference values and plot to be compared with values and plot you got everytime you change settings in the app.
Please note: part of the code is hidden just to save some space.

---

## Values and Graph for all cars in the dataset

```{r,echo=FALSE,results='hide'}
library(datasets)
library(UsingR)
library(graphics)
```

Average Weight:
```{r,echo=TRUE}
mean(mtcars$wt)
```

Average MPG:
```{r,echo=TRUE}
mean(mtcars$mpg)
```

```{r,echo=TRUE}
data$am <- as.factor(data$am)
mpgmean <- mean(mtcars$mpg)
wtmean <- mean(mtcars$wt)
plot(mtcars$wt, mtcars$mpg, col=mtcars$am, xlim=c(1,6), ylim=c(10,40), 
     ylab="MPG = Miles per Gallon", xlab="Weight (lbs/1000)", cex.main=1,
     main="MPG by Weight; Orange Lines =  mean weight and MPG for all cars",
     pch=16, cex=3)
abline(h = mpgmean, col="orange", lwd=4)
abline(v = wtmean, col="orange", lwd=4)
```

---

## Conclusions

By analyzing the different distributions on the plot of the cars corresponding to your chosen settings, and by matching
that with the averages of weights and MPGs values for the whole dataset (represented by the orange lines)
you can draw a first raw evaluation about the influence of engine size and trasmission type on the car size
and then on the car efficiency.
Remember that more efficient cars have a bigger MPG value, and so they are gathered into the two upper quadrants of the graph

You can see that the engine size has quite a direct influence on the weight and fuel consumption,
while the influence of transmission type is not to straight to be deciphered.
The reason of that is due to the fact that most of the automatic cars are heavier and with bigger engines with respect to cars with manual trasmission.
So, we'd need a bigger sample than that offered by mtcars dataset to be able to draw more definitive conclusions.